---
title: "Sardana_Module5-HW_files"
output:
  pdf_document: default
  word_document: default
date: "2023-02-13"
---
## Problem 1 Maximum Likelihood estimation 
1 a) Numerical Optimization given by Numerical_me
1 b) Analytical estimation given by Analytic_me
```{r}
# Maximum likelihood function
likelihood <- function(lam) prod(dexp(c(1.433, 0.524, 0.384, 4.515, 1.852, 0.429), rate = lam))

# Negative log-likelihood function
neglik <- function(lam)-sum(log(dexp(c(1.433, 0.524, 0.384, 4.515, 1.852, 0.429), rate = lam)))

# Maximum likelihood estimate numerical
Numerical_me <- optim(par=1,neglik)
Numerical_me$par


# Analytical maximum likelihood estimate
observations <- c(1.433, 0.524, 0.384, 4.515, 1.852, 0.429)
Analytic_me = 1/mean(observations)
Analytic_me


```

## Problem 2 
a) According to chi square distribution
If the population mean is m for a  random sample X1,X2,...Xn, 
point estimator of m is 98.6 

b) One sided 90% confidence interval Chi square distribution
```{r}
# sample size = 95
n <- 75
# degree of freedom = n-1 
df = n-1
# at low confidence interval 1-alpha = 0.90 , alpha = 0.01 and at alpha/2 = 0.05
# calculating chisquare at low confidence interval
qchisq(c(.05),df=74, lower.tail=FALSE)

```

## Problem3 
a) 
```{r}
# Load multtest package
library(multtest)
data(golub)
grep("Zyxin",golub.gnames[,2])
# Extracting the Zyxin gene data from golub at 2124 row
Zyxin <- golub[2124,]
# split the gene factor into 2 ALL and AML group
gene_factor <- factor(golub.cl,levels = 0:1,labels = c("ALL","AML"))

All_exp <- Zyxin[gene_factor == "ALL"]
AML_exp <- Zyxin[gene_factor == "AML"]

# To calculate the the bootstrap 95% CIs for the mean and variance of the ALL group
n <- length(All_exp)
nboot <- 1000
boot.xbar <- rep(NA,nboot)
boot.var <- rep(NA,nboot)

for (i in 1:nboot){
  data.star <- All_exp[sample(1:n,replace = TRUE)]
  boot.xbar[i] <- mean(data.star)
  boot.var[i]<- var(data.star)
} 

Mean_allCI <- quantile(boot.xbar, c(0.025, 0.975))
Var_allCI <- quantile(boot.var, c(0.025, 0.975))

# Print the mean and variance of ALL group 
print(sprintf("At bootstrap 95%% CIs , the mean of the ALL group: [%.3f, %.3f]", Mean_allCI[1], Mean_allCI[2]))

print(sprintf("At bootstrap 95%% CIs , the variance of the ALL group: [%.3f, %.3f]", Var_allCI[1], Var_allCI[2]))

# To calculate the the bootstrap 95% CIs for the mean and variance of the AML group

n <- length(AML_exp)
nboot <- 1000
boot.xbar <- rep(NA, nboot)
boot.var <- rep(NA, nboot)

for (i in 1:nboot) {
  data.star <- AML_exp[sample(1:n, replace=TRUE)]
  boot.xbar[i] <- mean(data.star)
  boot.var[i] <- var(data.star)
}

Mean_amlCI <- quantile(boot.xbar, c(0.025, 0.975))
Var_amlCI <- quantile(boot.var, c(0.025, 0.975))

# Print the mean and variance bootstrap CIs for the AML group
print(sprintf("At bootstrap 95%% CIs , the mean of the AML group: [%.3f, %.3f]", Mean_amlCI[1], Mean_amlCI[2]))

print(sprintf("At bootstrap 95%% CIs , the variance of the AML group: [%.3f,%.3f]", Var_amlCI[1], Var_amlCI[2]))


```
## Problem set 3 
(b)	to find the parametric 95% CIs for the mean and for the variance of the gene
# For AML group 

```{r}
mean_aml <- mean(AML_exp)
mean_aml_SE <- sd(AML_exp) / sqrt(length(AML_exp))
mean_aml_CI <- mean_aml + qnorm(c(0.025, 0.975)) * mean_aml_SE

# For the ALL group
mean_all <- mean(All_exp)
var_all <- var(All_exp)
df_all <- length(All_exp) - 1
var_all_CI <- qchisq(c(0.025, 0.975), df = df_all) * var_all / df_all

# For the AML group
var_aml <- var(AML_exp)
df_aml <- length(AML_exp) - 1
var_aml_CI <- qchisq(c(0.025, 0.975), df = df_aml) * var_aml / df_aml

# Results calculating the mean and variance of all and aml groups

#cat("ALL group:")
#cat("Mean:", mean_all, "CI:", mean_all_CI, "\n")
#cat("Variance:", var_all, "CI:", var_all_CI, "\n")

#cat("AML group:")
#cat("Mean:", mean_aml, "CI:", mean_aml_CI, "\n")
#cat("Variance:", var_aml, "CI:", var_aml_CI, "\n")

```

## Problem 3 set
(c)
To Find the bootstrap 95% CI for the median gene expression in aml and all groups separately.

```{r}
# Bootstrap median for ALL group
Zyxin<-golub[2124,]
n<-length(Zyxin)
nboot <- 1000
median_all <- numeric(nboot)
for (i in 1:nboot) {
  boot_all <- All_exp[sample(1:length(All_exp), replace = TRUE)]
  median_all[i] <- median(boot_all)
}

median_all_CI <- quantile(median_all, c(0.025, 0.975))

# Bootstrap median for AML group

median_aml <- numeric(nboot)
for (i in 1:nboot) {
  boot_aml <- AML_exp[sample(1:length(AML_exp), replace = TRUE)]
  median_aml[i] <- median(boot_aml)
}

median_aml_CI <- quantile(median_aml, c(0.025, 0.975))

cat(" The median gene expression for ALL group is:", median_all_CI, "\n")
cat("The median gene expression for AML group is:", median_aml_CI, "\n")


```
## Problem set 4
(a) and (b)

```{r}

nsim<-1000
MCsim<- function(nsim, lambda){
  cov1<-cov2<-rep(NA,nsim) 
  for (i in 1:nsim){
    x = rpois(50,lambda)
    xbar = mean(x)
    Xsd = sd(x)
    CI1<-c(xbar+(qt(0.05, 49)*sqrt(xbar/50)), xbar+qt(0.95, 49)*sqrt(xbar/50))
    CI2<-c((49*(Xsd^2)/qchisq(0.95, 49)), 49*(Xsd^2)/qchisq(0.05, 49))
    cov1[i]<-(CI1[1]<lambda)&(lambda<CI1[2])
    cov2[i]<-(CI2[1]<lambda)&(lambda<CI2[2])
    }
  print(paste("When lambda=", lambda, ": coverage for first CI is", mean(cov1), ", coverage for second CI is", mean(cov2), ".")) 
  }

# Used monte carlo simulations for nsim = 1000 runs at different parameter values
MCsim(nsim = 1000, lambda = 0.1)
MCsim(nsim = 1000, lambda = 1)
MCsim(nsim = 1000, lambda = 10)


```






