---
title: "**Sardana_Module3HW.Rmd**"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
date: "2023-01-31"
---

Problem 1 (a)
Given function is Poisson distribution 
1.# P(X=1)
```{r}

P <- dpois(1, lambda = 3)
P

```

Problem 1 b) # P(-3<X<5). 

```{r}

x_range <- c(-3:5)
sum(dpois(x_range, lambda = 3))

```

Problem2 
```{r}
n=5 
p = 1/7
```

Problem3

Binomial distribution #(n = 3, p = 0.25) 
# To compute 
a)	#P(Y ≤ 2)
```{r}
pbinom(2,size = 3, p = 0.25)
```

b)	E(Y) 
```{r}
EY <- 3*0.25
EY

```

c) Var(Y) 
```{r}
VarY <- 3 * 0.25 * (1-0.25)
VarY

```

Problem 4

Chi square distribution
a) P(2 < X < 5) 
```{r}
pchisq(5,df=5) - pchisq(2,df=5)

```

b) E(X) 

```{r}
a <- 5	
EX <- a
EX

```

c) Var(X)

```{r}
a <- 5
Var_X <- 2 * a
Var_X

```

d) Monte Carlo simulation
yes answer a) follows monte carlo stimulation 

```{r}
n <- 100000
df <- 5
chis_estimate <- rchisq(n,df)
chis_estimate <- sum(chis_estimate > 2 & chis_estimate < 5)/n
chis_estimate


```

Problem 5
Chisquare distribution
a) Compute E(Y) and Var (Y)

```{r}
E_X = 8

E_Y = 3*E_X-5
E_Y

Var_X <- 12
Var_Y <- 3^2*Var_X
Var_Y


```

b) No it does not follow a Chisquare distribution with degree of freedom m = 6
Variance should be 38 as Ey value is 19 and variance should be 2*19 = 38. 

Problem 6 

Normal distribution


a) Probability that a randomly chosen patient have the Zyxin gene expression values between 1 and 1.6



```{r}
mu <- 1.6
deviation <- 0.4
probability <- pnorm(1.6, mean = mu, sd = deviation) - pnorm(1, mean = mu, sd = deviation)
probability

```

b) Monte Carlo Simulation
# Estimate the probability as the proportion of samples between 1 and 1.6

```{r}
n <- 500000
mu <- 1.6
deviation <- 0.4

simulation <- rnorm(n, mean = mu, sd = deviation)

estimate_value <- mean(simulation >= 1 & simulation <= 1.6)
estimate_value

```

c)# Probability that exactly 2 out of 5 patients have the Zyxin gene expression values between 1 and 1.6
Binomial distribution
```{r}
#probability <- pnorm(1.6, mean = mu, sd = deviation) - pnorm(1, mean = mu, sd = deviation)
#probability
n <- 5
p <- probability
prob <- dbinom(2, size = n, prob = p)
prob

```


