---
title: Assignment2
output:
  pdf_document: default
  html_document: default
date: "2023-01-23"
---
Problem 1
Compute the gene means of golub data set

 Installed package multtest from BiocManager 

```{r,echo=FALSE,warning=FALSE, message = FALSE}
if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

BiocManager::install("multtest")
BiocManager::install("ALL")

library(multtest)
data(golub)
gol.fac <- factor(golub.cl, levels=0:1, labels = c("ALL","AML"))
```

Problem 1 
a) mean expression values for every gene among ALL patients


```{r}
meanALL <- apply(golub[,gol.fac=="ALL"], 1, mean)
# meanALL

```

b) mean expression values for every gene among AML patients
```{r}
meanAML <- apply(golub[,gol.fac=="AML"], 1, mean)
# meanAML
```

c) biological names of three genes with largest mean expression among AML patients

```{r}
meanALL <- apply(golub[,gol.fac=="ALL"], 1, mean)
o <- order(abs(meanALL), decreasing=TRUE)
print(golub.gnames[o[1:3],2])
```

d) biological names of three genes with largest mean expression value among AML patients

```{r}
meanAML <- apply(golub[,gol.fac == "AML"], 1,mean)
o <- order(abs(meanAML),decreasing = TRUE)
print(golub.gnames[o[1:3],2])
```

Problem 2 

a) To save the expression values of first five genes for AML patients in a csv
```{r,eval=FALSE}
firstFive <- golub[1:5,gol.fac == "AML"]
write.csv(firstFive,"AML5.csv")

```

b) To save the expression values of first five genes for ALL patients in a text
```{r,eval=FALSE}
first_five_ALL <- golub[1:5,gol.fac =="ALL"]
write.table(first_five_ALL,"ALL5.txt")
```

c) Compute the standard deviation of expression values on 1st patient of 100th to 200th genes
```{r}

patient1 <- (golub[100:200,1])
standardDeviation <- sd(patient1)
print(standardDeviation)
```

d) To compute the standard deviation of every gene across all the patients
```{r}
standard_deviation <- apply(golub,1,sd)
print(length(which(standard_deviation>1)))

```

e) A scatter plot of the 101th gene expressions against the 102th gene expressions

```{r}
gene_exp = golub[101,]
 
 
gene_exp2 = golub[102,]

 
name1 = golub.gnames[101,2]
name1
 
name2 = golub.gnames[102,2]
name2
 
plot(gene_exp,gene_exp2,xlab =name1,ylab = name2, main = "scatterplot of gene expression 101th against 102th")

```

Problem 3

Work with the ALL data set, Load the ALL data from the ALL library and use str and openVignette() for a further orientation.

```{r}
library("ALL")
str("ALL")
# openVignette("ALL")

```

a)  Produce one histogram of gene expression using exprs(ALL[,ALL$BT=="B1"] to extract the gene expression from patients in disease stage B1

```{r,eval = FALSE}

gene_expression_B1 <- exprs(ALL[, ALL$BT == "B1"])
 hist(gene_expression_B1, xlab = "Gene Expression", ylab = "Frequency", main = "Histogram plot of Gene Expressions in Disease Stage B1")

```

b) mean gene expressions for every gene over these B1 patients.

```{r, eval = FALSE}
meanB1 <- apply(exprs(ALL[,ALL$BT=="B1"]),1, mean)
head(meanB1)
```

c) gene identifiers of the three genes with the largest mean
```{r,eval = FALSE}

meanB1 <- apply(exprs(ALL[,ALL$BT=="B1"]),1, mean)
sorted <- sort(meanB1, decreasing=TRUE)
print(sorted[1:3])
```


Problem 4

To produce 2 overlaid scatterplots Height vs Girth, Volume vs Girth.

```{r}
 plot(trees$Girth, trees$Height ,pch = "+", xlab = "Girth", main = "Overlaid scatter plot of Girth vs height, Girth vs volume", ylim = c(0, max(trees$Height, trees$Volume)))
 points(trees$Girth, trees$Height, pch = "+", col = "blue")
 points(trees$Girth, trees$Volume, pch = "o", col = "red")
 legend("topright", c("Height", "Volume"), col = c("blue", "red"), pch = c("+", "o"))
 
```
