---
title: "**Assignment1**"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
date: "2023-01-16"
---

Problem 1 (a)
This type of vector is numeric
used type of function to predict the class of the object
```{r}

vec <-c(5,TRUE) 
typeof(vec)

```
Problem 1 (b) To calculate the expression of given vectors x+y
```{r}

x <- 1:4
y <- 1:2
x+y

```
Problem 1 (c) the R command c(1,2) %*% t(c(1,2)) 
The code returns a two by two matrix

```{r}

c(1,2) %*% t(c(1,2))

```
Problem 1 (d) Running the provided R statement gives the output as 10

```{r}

f <- function(x) { 
  g <- function(y) { 
    y+z
   }
  z<-4
  x+g(x)
}

z<-15
f(3)

```
Problem 2 Use R to calculate the provided expression in question

```{r}

sum <- 0
for(x in 1:1000){
  
    sum = sum +x^2
  
}
print(sum)


```
Question 3 
(a)	Create a vector named “age” to represent reasonable age of a person
```{r}

age_years <- c(10,15,20,25,30,35,40,45,50,55)
age_years

```
(b)	Multiply each person’s age by 12 (to convert into months). 
```{r}

person_age_months <- age_years*12
person_age_months

```
c)	Find the sum of ages of all these people. 

```{r}

sum(person_age_months)

``` 
d)	Find the age of the youngest person
```{r}

min(person_age_months)

``` 
e)	Find the age of the oldest person.

```{r}

max(person_age_months)

``` 
f)	Find the square root of the age of each person. (Not sure what this means, but who cares?) (this also should be a vector)

```{r}

sqrt(person_age_months)

``` 
Question 4 
(g)	Create a vector X of length 30, with the kth element in X = 3k, for k=1…30. Print out the values of X.

```{r}

X <- seq(1,30 ,by =3)
x1 <- 3*X
x1

``` 
h)	Create a vector Y of length 30, with all elements in Y equal to 0. Print out the values of Y.
```{r}

Y <- rep(0,30)
Y

``` 
i)	Using a “for” loop, reassigns the value of the k-th element in Y, for k = 1…30. When k less than 20, the kth element of Y is reassigned as the sine of (2k).  When the k greater than  20,  the kth element of Y is reassigned as the value of integral  .  (You may want to use $value at the end of the line to get the integration with R clean out unwanted values)

```{r}

Y <- rep(0,30)
for (k in 1:30){
  if (k<20){
    Y[k] = sin(2*k)
  } else if (k>= 20){
    Y[k] = integrate(sqrt, lower = 0, upper = k)$value
  }
}
print(Y)

``` 

